---
title: "WUP_04"
author: "Destiny Calvillo"
date: "Summer 2021"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

```{r}
gssdata=read_csv("GSScsv.csv")
```

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(gssdata,aes(x=DEGREE,y=WORDSUM))
```
## Now add in boxplot

```{r}
basicCN + geom_boxplot()
```

It looks like the box of LT High School is completely different than both Graduate and Junior College. If we were to slide the box under the two, it would be completely underneath them. This shows that the middle 50 percent of the High School box is completely lower than Graduate and Junior College, which causes me to suspect that they will be different.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=gssdata)
```

The High School, Bachelor, and Junior College means are all within the same area as they are really close in number to each other. As for LT High School, it is a bit lower with a mean of 7.75. The Graduate mean is also higher in scale eith a mean of approximately 10.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=gssdata)
summary.aov(model1)
```


```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 13.2, 2, 2))
plot(TukeyHSD(model1),las=2)
```

The only things that are different are the LT High School-Graduate and the High School-Graduate levels of degrees. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

We first found out that it was a numerical and categorical problem with more than two categories. So, we got our helper file in order to get the one-way anova tools. After running th first test we noticed a significant result, which resulted in using the Tukey HSD tool to find out what was different from what. We resulted in agreeing with the alternative hypothesis, which is that there is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE.
